Moment Constraints Make Learning Latent-Variable Models Easier

### 2014-01-27 doc/icml2014/background.tex
Updated so many things to figures

\section{Background}
\label{sec:background}

\paragraph{Bottlenecks: Three-view mixture models}

@@

\todo{Related work on spectral methods and measurements.}

\paragraph{Constrained EM}

Posterior regularization
Measurements
Generalized expectation criterion

Kuzman Ganchev.*

>>>>

\section{Setup}
\label{sec:setup}

Let $\sG$ be a discrete graphical model with variables $V$, of which
  $O \subseteq V$ be observed and $H = V \setminus O$ be hidden.
Let $\Pa(x) \subset V$ be the parents of a variable $x$ in a directed
  model, and $\sN(x) \subset V$ be the neighbours of $x$ in an undirected
  model.
We will use $o_1, o_2, \cdots \in O$ to denote observed variables and
  $h_1, h_2, \cdots \in H$ to denote hidden variables.
For ease of exposition, we assume that the domain of all the observed
  variables is $[D]$ and the domain of all the hidden variables is
  $[K]$.

Define $\otimes$ to be the tensor product. In matrix representation, it
  is the Kronecker product.\expand

\paragraph{Problem statement}

% Statement
This paper focuses on the problem of parameter estimation:
We are given $n$ i.i.d.~examples of the observed variables $D = (x^{(1)}, \dots, x^{(n)})$
where each $x^{(i)} \sim p_{\theta^*}$ for some true parameters $\theta^*$.
Our goal is to produce a parameter estimate $\hat\theta$ that approximates $\theta^*$.

% Maximum likelihood
The standard estimation procedure is maximum (marginal) likelihood,
  \begin{align*}
    \sL &\eqdef \max_{\theta \in \Re^d} \E_{x} \log p_\theta(x) \\
        &=      \max_{\theta \in \Re^d} \E_{x} \log \sum_z p_\theta(z),
  \end{align*}
  which is statistically efficient but computationally intractable
  because we must marginalize over latent variables $z$.
In practice, one uses gradient-based optimization procedures (e.g., EM
  or L-BFGS) on the marginal likelihood, which can get stuck in local
  optima.

\paragraph{Estimating three-view mixture models}

@@

In general, we call hidden variables with this property ``bottlenecks'',
  and define them as follows.
\begin{definition}(Bottleneck)
  A hidden variable $h$ is said to be a bottleneck if there exists at
  least three conditionally independent observed variables $o_1, o_2,
  o_3$. Let $\sB(h)$ be the set of conditionally independent observed
  variables for $h$.
\end{definition}

\begin{assumption}
  We assume that $M_1, M_2, M_3$ are each full rank, and that $\pi \succ
  0$\reword.
\end{assumption}

### 2014-01-27 doc/icml2014/intro.tex
Updated so many things to figures

% 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
Recently, unsupervised learning techniques based on the method of moments and
spectral decomposition have offered a refreshing and promising perspective on
this learning problem \citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
These methods exploit the linear algebraic properties of the model and
factorize the moments into parameters, providing strong theoretical guarantees.
However, these methods are not as universally applicable as EM.

% 4. State what we do: exploit moment constraints to make the problem easier.
In this work, we exploit the spectral method to learn moment constraints
on the observed variables and show how these constraints can greatly
simplify the learning problem.
% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals.
The key idea in
\citet{anandkumar12moments,anandkumar13tensor} is to study the conditional
independence factorization structure implied by the model in the form of 
three-view bottlenecks. These bottlenecks which can be factorized into
moments involving the latent variables\vague.

% 6. Provide some intuitive examples - Can I haz HMMs?
Consider a hidden Markov model, with initial probabilities $\pi$,
  transitions $T$ and emissions $O$. 
While \citet{hsu09spectral,anandkumar12moments} provide a consistent estimator
  for the parameters of this model that inverts $O$ to recover $T$ and
  $\pi$. We would like to avoid this potentially sensitive procedure via optimization.
The log-likelihood of this model, 
  \begin{align}
    \sL(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h} \pi(h_1) O(x_1 | h_1) T(h_2 | h_1) O(x_2 | h_2) \cdots ),
  \end{align}
  is not convex, even if we knew any one of the parameters.
However, intuitively, if we knew $O$ and found that each symbol was
  produced by exactly one state, learning $T$ and $\pi$ could be trivial
  accomplished via counting. 
In general, $O$ still informs us about the states in the observation
  sequence. If we considered instead the piecewise likelihood, 
  \begin{align}
    \sL_{\textrm{p}}(\pi, T, O) &= \sum_{x\in\sD} \log( \sum_{h_1, h_2} \underbrace{\pi(h_1) T(h_2 | h_1)}_{P(h_1,h_2)} O(x_1 | h_1) O(x_2 | h_2) ),
  \end{align}
  then we observe that the likelihood is convex in the marginal
  distribution $P(h_1,h_2)$. This allows us to recover $P$, from which $T$
  can be easily reconstructed.

% 7. Now outline the rest of our results.
In \sectionref{piecewise} we generalize this observation to a directed grid
  model and a directed tree model and show that the {\em piecewise
  likelihood} is strictly convex, providing consistent parameter
  estimators for these models (which is new)\reword.
The approach can be extended to log-linear
  models (\sectionref{log-linear}), where the problem remains non-convex,
  however, we expect a reduction in the number of local optima. 
In this regime, we show empirically that constraining the moments of the
  model via posterior-regularization allows us to learn better models.

>>>>

% 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
Recently, unsupervised learning techniques based on the method of moments and
spectral decomposition have offered a refreshing and promising perspective on
this learning problem \citep{hsu09spectral,anandkumar11tree,anandkumar12moments,anandkumar12lda,hsu12identifiability,balle11transducer,balle12automata}.
These methods exploit the linear algebraic properties of the model and
factorize the moments into parameters, providing strong theoretical guarantees.
However, these methods are not as universally applicable as EM, applying to a limited set of models.

\begin{figure*}
  \centering
  \input{figures/approach.tikz}
  \caption{Overview of approach}
\end{figure*}

% 4. State what we do: exploit moment constraints to make the problem easier.
In this work, we exploit the spectral method to learn moment constraints
on the observed variables and show how these constraints lead to consistent parameter estimates for a broad class of graphical models.

% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals.
Our approach is illustrated in \figureref{approach}. 
First, we identify conditionally independent observed variables $x_1,
  x_2, x_3$ for each hidden variable $h$ (bottlenecks). 
We use the tensor factorization algorithm of
  \citet{anandkumar12moments,anandkumar13tensor} to identify the
  conditional moments.
Given these factors, we show that for every clique of hidden variables
  $\sC$, the piecewise likelihood of the observed variables is convex in
  the marginal distribution. 
This guarantees that we can recover the true marginal distributions
  using EM; finally, with some post-processing, we are able to
  consistently recover the true parameters of the model.
In \sectionref{piecewise}, we detail our algorithm and describe the
  settings and technical assumptions in which this approach is guaranteed
  to work.
As a particular example, we are able to recover parameters for the
  grid model depicted in \figureref{approach} under the condition that
  $O$ is full-rank.

For undirected log-linear latent variable models, this approach allows us to 
  recover the marginals for each clique, given which the learning
  problem is convex (\sectionref{undirected}).
Finally, our approach gracefully extends to more general model families
  in which we are unable to recover the marginal distributions for all
  cliques; in such scenarios, we use posterior regularization to constrain
  EM. 
Empirically, we find that this approach leads to better solutions
  (\sectionref{experiments}).
